\section{Conclusion}
This chapter proposes a semi-supervised recursive multivariate piecewise motif mining approach 
to electricity disaggregation. 
%Before unsupervised disaggregation, 
We use data from a period of time to find the features of individual devices, 
including the power level and standard deviation.
Based on these features, 
we recursively utilize multivariate motif mining to uncover devices 
that draw power from both phases equally or unequally, 
then separate the devices from each phase using motif mining. 
Devices that use a large amount of power are removed from the two phases in the first few rounds, 
which brings the benefit of decreasing the noise caused by these devices 
during the process of disaggregating from single phase. 
Therefore more devices with smaller power consumption are separated 
using the piecewise motif mining from a single phase. 
In addition, this piecewise motif mining approach can identify
continuously variable loads such as outdoor heating. 
Furthermore, when we apply motif mining approach to water disaggregation, 
it can separate water use ends that have steady water usage, such as a shower, 
but cannot disaggregate those water use ends which consume water variably for the whole 
cycle, like a toilet. 

In the future, 
for energy disaggregation we will use more features from the aggregated data from the multiple phases, 
such as the startup shape of individual devices.
For water disaggregation, we will explore how to integrate the 
dynamic time warping with multivariate piecewise motif mining. 
